---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  out.width = "100%"
)
```

# regmdc

<!-- badges: start -->
<!-- badges: end -->

regmdc is an R package for fitting nonparametric regression models with mixed 
derivative constraints. Available estimation methods are entirely monotonic 
regression, Hardy—Krause variation denoising, totally convex regression, MARS 
via LASSO, and their generalizations. Entirely monotonic regression and 
Hardy—Krause variation denoising are multivariate generalizations of isotonic 
regression and total variation denoising, introduced in Fang et al. (2021). 
MARS via LASSO is a LASSO variant of the usual MARS method, introduced in 
Ki et al. (2021). It can be thought of not only as a multivariate 
generalization of locally adaptive regression splines but also as second-order 
Hardy—Krause variation denoising. Totally convex regression is a multivariate 
extension of univariate convex regression based on total convexity (see, e.g., 
Gal (2008)). A document for totally convex regression is now in preparation. 

regmdc mainly consists of the following two generic functions:

* `get_lasso_problem_soln()`
* `compute_fit()`

Given an estimation method, `get_lasso_problem_soln()` finds the solution to the 
corresponding constrained LASSO problem. For details on the LASSO problem of 
each estimation method, see, for example, Section 3 of Fang et al. (2021) (for 
entirely monotonic regression and Hardy—Krause variation denoising) and 
Section 3 and 6 of Ki et al. (2021) (for MARS via LASSO). There are some other 
ongoing research and working papers, and they would be available in the future.

Given the solution to the LASSO problem obtained from `get_lasso_problem_soln()`, 
`compute_fit()` computes the fitted values of an estimation method.


## Installation

You first need to install [MOSEK](https://docs.mosek.com/latest/install/installation.html), 
a software package for optimization, and then [Rmosek](https://docs.mosek.com/latest/rmosek/install-interface.html), 
the MOSEK interface in R.

Once they are properly installed, you can install our package by running in R 
the following commands:

```r
# install.packages("devtools")
devtools::install_github("DohyeongKi/regmdc")
```


## Examples

Here are some examples showing how regmdc can be used. Please refer to the 
document of each function in the package for more details.

```{r em_example, eval = FALSE}
################################# 
# Entirely monotonic regression #
#################################
library(regmdc)

fstar <- function(x) {x[1] + x[2]}  # the underlying function

X_design <- expand.grid(rep(list(seq(0, 9.0/10, length.out = 10L)), 2L))  # a design matrix
# Compute the values of f* at the design points
theta <- apply(X_design, MARGIN = 1L, FUN = fstar)

sigma <- 1  # standard Gaussian noise
y <- theta + sigma * rnorm(nrow(X_design))  # an observation vector

# Obtain the solution to the corresponding constrained LASSO problem
raw_soln <- get_lasso_problem_soln(X_design, y, s = 1L, method = "em")$solution

# Remove the zero components from the solution
threshold <- 1e-4  # a threshold to determine whether each component of the solution is zero
is_nonzero <- (abs(raw_soln) > threshold)
processed_soln <- as.numeric(is_nonzero) * raw_soln
compressed_soln <- processed_soln[processed_soln != 0]

X_eval <- expand.grid(rep(list(seq(0, 1, length.out = 51L)), 2L))  # an evaluation matrix
# Compute the fitted values at the evaluation points
compute_fit(X_eval, X_design, s = 1L, method = "em", compressed_soln, is_nonzero)
```

```{r hk_example, eval = FALSE}
#################################### 
# Hardy—Krause variation denoising #
####################################
library(regmdc)

fstar <- function(x) {x[1] - x[2]}  # the underlying function

X_design <- expand.grid(rep(list(seq(0, 9.0/10, length.out = 10L)), 2L))  # a design matrix
# Compute the values of f* at the design points
theta <- apply(X_design, MARGIN = 1L, FUN = fstar)

sigma <- 1  # standard Gaussian noise
y <- theta + sigma * rnorm(nrow(X_design))  # an observation vector

# Obtain the solution to the corresponding constrained LASSO problem
raw_soln <- get_lasso_problem_soln(X_design, y, s = 2L, method = "hk", V = 2)$solution

# Remove the zero components from the solution
threshold <- 1e-4  # a threshold to determine whether each component of the solution is zero
is_nonzero <- (abs(raw_soln) > threshold)
processed_soln <- as.numeric(is_nonzero) * raw_soln
compressed_soln <- processed_soln[processed_soln != 0]

X_eval <- expand.grid(rep(list(seq(0, 1, length.out = 51L)), 2L))  # an evaluation matrix
# Compute the fitted values at the evaluation points
compute_fit(X_eval, X_design, s = 2L, method = "hk", compressed_soln, is_nonzero)
```

```{r emhk_example, eval = FALSE}
######################################################################################## 
# Generalization of entirely monotonic regression and Hardy—Krause variation denoising #
########################################################################################
library(regmdc)

fstar <- function(x) {x[1] - x[2] + x[1] * x[2]}  # the underlying function

X_design <- expand.grid(rep(list(seq(0, 9.0/10, length.out = 10L)), 2L))  # a design matrix
# Compute the values of f* at the design points
theta <- apply(X_design, MARGIN = 1L, FUN = fstar)

sigma <- 1  # standard Gaussian noise
y <- theta + sigma * rnorm(nrow(X_design))  # an observation vector

# Obtain the solution to the corresponding constrained LASSO problem
raw_soln <- get_lasso_problem_soln(X_design, y, s = 2L, method = "emhk", V = 1,
                                   constrained_interactions = c('1-2'),
                                   positive_interactions = c('1'),
                                   negative_interactions = c('2'))$solution

# Remove the zero components from the solution
threshold <- 1e-4  # a threshold to determine whether each component of the solution is zero
is_nonzero <- (abs(raw_soln) > threshold)
processed_soln <- as.numeric(is_nonzero) * raw_soln
compressed_soln <- processed_soln[processed_soln != 0]

X_eval <- expand.grid(rep(list(seq(0, 1, length.out = 51L)), 2L))  # an evaluation matrix
# Compute the fitted values at the evaluation points
compute_fit(X_eval, X_design, s = 2L, method = "emhk", compressed_soln, is_nonzero)
```


## References

[1] Ki, D., Fang, B., and Guntuboyina, A. (2021). MARS via LASSO. 
https://arxiv.org/abs/2111.11694.

[2] Fang, B., Guntuboyina, A., and Sen, B. (2021). Multivariate extensions of 
isotonic regression and total variation denoising via entire monotonicity and 
Hardy—Krause variation. *The Annals of Statistics*, **49**(2), 769-792.

[3] Gal, S. G. (2008). *Shape-Preserving Approximation by Real and Complex Polynomials*. 
Birkhäuser, Boston.
